Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-03 demystify.tex
updates

>>>>

\section{I am a member of an underrepresented group in CS}\label{sec:urm}
One question that often comes up is if you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group. This question arises often for international (and even domestic) students. The main worry is that they might get discriminated against, as often happen in their countries or personal experiences. %, and (ii) mentioning this might make them look like they are asking for special treatment.

In my personal opinion, the short answer is \emph{yes}, it \emph{does} help, especially if you can articulate your experiences and how you can contribute to diversity and inclusion in CS. Most universities have a strong commitment to diversity and inclusion and want to  
\emph{increase the number of URM students} in their programs (e.g., GMU is consistently ranked as one of the most diverse universities in the US). Moreover, many universities and industrial organizations also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).

It is understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a prof who might be biased against URM students.  However, if a prof. has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior). In my experience, it is the opposite: many profs. are actively looking for URM students to work with, and mentioning this in your application can help you stand out.

In addition to being aware of the importance of diversity and inclusion in CS and want to contribute to it, profs. have incentives to work with URM students. For example, professors often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}). If in doubt, seek advice from your close professors and collaborators about this.
